Paper on the carbonate chemistry at Ny-Ålesund

### 2020-04-22 paper.Rmd
completedness

```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
#R --quiet -e 'install.packages("plotly", type="binary", repos="http://cran.us.r-project.org")'
Sys.setlocale("LC_ALL", "en_US.UTF-8")
Sys.setenv(TZ='UTC') # on utilise UTC
rm(list = ls())
library(tidyverse)
library(seacarb)
library(gridExtra)
library(lubridate)
library(viridis)
require("knitr")
library(kableExtra)
library("lmodel2")
library(scales)
library(plotly)
library(htmlwidgets)
library(tsibble)
if (!require("directlabels")) install.packages("directlabels")
library("directlabels")
if (!require("pangaear")) install.packages("pangaea\r")
library("pangaear")
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = normalizePath("/Users/gattuso/not_synced_git/AWIPEV-CO2-git"))

@@

######## function to calculate 3 quantiles
qq <- function (data, digits) { 
  as.numeric(formatC(quantile(x = data, probs = c(0.025, 0.50, 0.975), na.rm = TRUE),
                     format = "f", digits = digits))
}

@@

```{r read and clean data, echo=FALSE, warning=FALSE, message=FALSE}
d_all <- as_tibble(readRDS(file = paste0(path_data, "fb_data/d_all.rds"))) 
d <- d_all %>% 
  #dplyr::select
  #dplyr::filter(sal_fb > 28) %>% # remove salinity data lower than 28
  dplyr::mutate(year_month = format(datetime, '%Y-%m')) %>%
  dplyr::arrange(datetime) %>%
  dplyr::mutate(depth = as.numeric(str_remove(depth, pattern= " m"))) #depth is a string
```

@@

## Temperature vs salinity
We explore the use of a TS diagram to investigate changes in water sources. Monthly slopes of the T vs S plots are calculated. It seems that slopes are mostly not significant in winter and that they can be both positive and negative in summer.

@@

## Consistency: pCO2

>>>>

```{r set-up, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
#R --quiet -e 'install.packages("plotly", type="binary", repos="http://cran.us.r-project.org")'
Sys.setlocale("LC_ALL", "en_US.UTF-8")
Sys.setenv(TZ='UTC') # on utilise UTC
rm(list = ls())
library(tidyverse)
library(seacarb)
library(gridExtra)
library(lubridate)
library(viridis)
require("knitr")
library(kableExtra)
library("lmodel2")
library(scales)
library(plotly)
library(htmlwidgets)
library(tsibble)
if (!require("directlabels")) install.packages("directlabels")
library("directlabels")
if (!require("pangaear")) install.packages("pangaea\r")
library("pangaear")
if (!require("naniar")) install.packages("naniar")
library("naniar")
if (!require("visdat")) install.packages("visdat")
library("visdat")
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = normalizePath("/Users/gattuso/not_synced_git/AWIPEV-CO2-git"))

@@

######## function to calculate min, max 3 quantiles
qq <- function (data, digits) { 
  as.numeric(formatC(quantile(x = data, probs = c(0, 0.025, 0.50, 0.975, 1), na.rm = TRUE),
                     format = "f", digits = digits))
}

@@

```{r read and clean data, echo=FALSE, warning=FALSE, message=FALSE}
# d_all <- readRDS(file = paste0(path_data, "fb_awipev-co2_server/ny-alesund/data/NRT_data/all_nydata_hour_copy.rds"))
# d <- d_all %>% 
#   as_tsibble(index = datetime) %>% 
#   dplyr::rename(depth = pressure_insitu_ctd) %>% 
#   dplyr::mutate(year_month = format(datetime, '%Y-%m'))
# count_gaps(d, .full = TRUE)

d_all <- readRDS(file = paste0(path_data, "fb_data/d_all.rds"))
d <- d_all %>% 
  #dplyr::select
  #dplyr::filter(sal_fb > 28) %>% # remove salinity data lower than 28
  as_tsibble(index = datetime) %>% 
  dplyr::arrange(datetime) %>%
  dplyr::mutate(depth = as.numeric(str_remove(depth, pattern= " m"))) %>% #depth is a string
  fill_gaps() %>%
  dplyr::mutate(year_month = format(datetime, '%Y-%m'),
                quarter = quarter(datetime)) %>%
  replace(., is.na(.), NA) # replace NaN by NA everywhere
tmp <- d %>% 
  as_tibble() %>% 
  dplyr::select(pH_s_seafet_temp_ctd, par_air_filtered, pco2_contros, pH_s_seafet_temp_ctd, starts_with("sal_"),
                starts_with("temp_"), at_calc, year_month, quarter)
tmp visdat::vis_miss(tmp)
gg_miss_var(tmp, facet = quarter)
```

```{r calculate missing, echo=FALSE, warning=FALSE, message=FALSE}

```
# {.tabset .tabset-fade .tabset-pills}

@@

## TvsS
We explore the use of a TS diagram to investigate changes in water sources. Monthly slopes of the T vs S plots are calculated. It seems that slopes are mostly not significant in winter and that they can be both positive and negative in summer.

@@

## Sensitivity
We look at how sensitive pH are to errors in salinity (0.1 to 0.3) and temperature (-0.3 to 0.3). delta_pH is the unit difference between the starting value of pH (8 at S=34 and T=5) and the value with errors in S and T. pCO2 is measured and not sensitive, I think to errors in temperature.

```{r sensitivity, echo=FALSE, warning=FALSE, message=FALSE}
err <- as_tibble(
  expand.grid(sal = 34+seq(0, 0.3, by = 0.1), 
              temp = 5+round(seq(-0.3, 0.3, by = 0.1), digits=1))) %>%
    dplyr::arrange(sal)
err$pH <- pHinsi(pH=8,ALK=2.4e-3,Tinsi=err$temp,Tlab=5,Pinsi=1,S=err$sal,Pt=0,Sit=0)
err$delta_pH <- round(8 - err$pH, digits = 3)
knitr::kable(err, caption = "Errors in pH related to the uncertainty in salinity and temperature.") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
saveRDS(err, file=paste0(path_data, "error.rds"))
```
## Consistency: pCO2
